Analysis and Figures for Gallego et al, 2020

### 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
Waiting for Wilcoxon

 Each biological cluster (colored hulls, Fig. 2.A) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig. 2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf. promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig. 2C). Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where community 3 is the most likely community -- to occur more than two-thirds of the time (68\%). {\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?}  \\

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

Diatoms show the steepest richness decline under future conditions (Figure 3); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Thalassionema}, \textit{Navicula}, \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Ditylum}) body forms. These declines in diatom richness were more accentuated at lower pH values, both in the warm and colder seasons. Other taxa likely to find less-suitable habitat in the future include \textit{Karlodinium} (dinoflagellate; San Juan Island), and \textit{Phaeocystis} and red-tide-associated \textit{Chattonella} (haptophyte and raphidophyte, respectively; Hood Canal).

Likely winners under future conditions are more widely scattered among higher taxonomic groups. The coccolithophore \textit{E. huxleyi} and the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas. Among others, \textit{Chaetoceros} (diatom) and two small hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in the warmer Hood Canal. See Supporting Information for a complete list of taxon-specific projections.

Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present (ANOVA; taxon richness within phyla by year, p $=$ 0.3). Diatoms, however, show consistent and significant declines in richness (Wilcoxon p $< 10^{-15}$).

@@

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-27.Other.sel.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

>>>>

 Each biological cluster (colored hulls, Fig. 2A) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig. 2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf. promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig. 2C). Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ {\color{red} 7.9  \st{8.2}} -- where community 3 is the most likely community -- to occur more than {\color{red} four-fifths \st{two-thirds}} of the time ({\color{red} 83\%  \st{68\%}}). {\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?} RYAN: It is a tricky sentence as we no longer estimate seasons independently. Cluster 3 is now mode prevalent on pH < 7.8 and Temperature > 12.5; and that condition appears now 83 \% of the time in our simulation.   \\

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-09-03.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

Diatoms show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., {\color{red} \st{\textit{Thalassionema}, \textit{Navicula},}} \textit{Nitzschia} ) and centric (e.g., \textit{Coscinodiscus},  {\color{red} \textit{Stephanodiscus}}) body forms. These declines in diatom richness were more accentuated at lower pH values{\color{red} and higher temperatures}. Other taxa likely to find less-suitable habitat in the future include the dinoflagellates {\color{red}\textit{Karlodinium} and \textit{Gymnodiniaceae} sp.\st{ and red-tide-associated \textit{Chattonella} (haptophyte and raphidophyte, respectively; Hood Canal).} }

Likely winners under future conditions are more widely scattered among higher taxonomic groups. The haptophytes \textit{E. huxleyi} and \textit{Chrysochromulina} sp. the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas. Among others, \textit{Chaetoceros} (diatom) and {\color{red} many} hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in {\color{red} both study areas}. See Supporting Information for a complete list of taxon-specific projections.
{\color{red}RYAN - I finished here today, will talk to you ab Wilcoxon} 
Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present (ANOVA; taxon richness within phyla by year, p $=$ 0.3). Diatoms, however, show consistent and significant declines in richness (Wilcoxon p $< 10^{-15}$).

@@

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-09-03.Other.sel.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

### 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
Waiting for Wilcoxon

```{r}
taxa.to.follow %>% 
  filter (p<0.01, average > 0.01) %>%  
  arrange(desc(average)) %>% 
  group_by(Area, Comparison) %>% 
    slice(1:5) %>%

  transmute(Area, Comparison, taxa = value, diff = avb - ava, average = case_when(ava > avb ~ average, TRUE ~ -average), sd) %>% 
  left_join(phylum.info %>% select(phylum,taxa)) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
  mutate(label2 = paste(phylum, label, sep = " | ")) %>% 
  mutate(Comparison = fct_relabel(Comparison, ~ str_replace(.,"_", " vs "))) %>% 
  nest(-Area) %>% 
  mutate(plot = map(data, function(.x){
    lims = c(-max(abs(.x$average) + (.x$sd)), max(abs(.x$average) + (.x$sd)))
    ggplot(.x, aes(y = average,
             x = fct_reorder(label2, average))) +
  geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
      geom_errorbar(aes(ymin = average -sd, ymax = average + sd ), width = 0.25) +
  geom_hline(yintercept = 0, color = "black")+
  labs (x = "taxa",
        y = "Contribution to dissimilarity")+
  coord_flip(ylim=lims) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  facet_wrap(~Comparison) +
  guides(fill = "none") +

@@

pmap(.l = list(Suppl.Material.plots.final, paths),
     .f = function(a,b) a +  ggsave(filename = here("Figures","Supplementary", b),
                                    device = "png",
                                    units = "in",
                                    dpi = "retina",
                                    width = 20,
                                    height = 12))

@@

>>>>

And keep some that are more abundant on each cluster too

```{r}
taxa.to.follow %>% 
  filter (average > 0.01) %>%  
  arrange(desc(average)) %>% 
  group_by(Area, Comparison, ava > avb) %>% 
    slice(1:5) %>%

  transmute(Area, Comparison, taxa = value, diff = avb - ava, average = case_when(ava > avb ~ average, TRUE ~ -average), sd) %>% 
  left_join(phylum.info %>% select(phylum,taxa)) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
  mutate(label2 = paste(phylum, label, sep = " | ")) %>% 
  mutate(Comparison = fct_relabel(Comparison, ~ str_replace(.,"_", " vs "))) %>% 
  nest(-Area) %>% 
  mutate(plot = map(data, function(.x){
    lims = c(-max(abs(.x$average) + (.x$sd)), max(abs(.x$average) + (.x$sd)))
    ggplot(.x, aes(y = average,
             x = fct_reorder(label2, phylum))) +
  geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
      geom_errorbar(aes(ymin = average -sd, ymax = average + sd ), width = 0.25) +
  geom_hline(yintercept = 0, color = "black")+
  labs (x = "taxa",
        y = "Contribution to dissimilarity")+
  coord_flip(ylim=lims) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  facet_wrap(~Comparison) +
  guides(fill = "none") +

@@

pmap(.l = list(Suppl.Material.plots.final, paths),
     .f = function(a,b) a + ggsave(plot = a,
                               filename = here("Figures","Supplementary", b),
                                    device = "png",
                                    units = "in",
                                    dpi = "retina",
                                    width = 20,
                                    height = 12))

@@

#### Bonus track 2: changes in richness, tested

```{r}
input.for.left %>% 
  nest(-phylum) %>% 
  #  slice(1) %>%
    mutate(data = map (data, ~ .x %>% 
                         )) %>% unnest(data)
                         pivot_longer(cols = c(-taxa, -sim.row),
                                      names_to = "draw.row",
                                      values_to = "presence")  %>% 
                         group_by(draw.row, sim.row) %>% 
                         summarise(tot = sum(presence)) %>% 
                         group_by(sim.row) %>% 
                         summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)) %>% 
  left_join(all.sims.for.projection %>%  
              rownames_to_column("sim.row")) ))%>%

   mutate(lm = map(data, ~ lm(Richness ~ Year, data = .x)),
         lm.results = map (lm, broom::tidy))  -> results.lm

### Using a wilcox test to see if there is a difference between Years

results.lm %>% 
  mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ),
         tidy.w   = map (wilcoxon, broom::tidy)) -> results.lm
